A Two-Infall Star Formation History with Outflows and Radial Migration

### 2025-02-18 src/tex/ms.tex

We increased $y_{\rm Fe}^{\rm Ia}$ by a factor of 30\% and 10\% for the {\tt yZ1} and {\tt yZ2} yield sets, respectively, to ensure our models could reach ${\rm [O/Fe]}\approx0.0$ by the end time. The final row of Table \ref{tab:yields} reports the integrated SN Ia rate $N_{\rm Ia}/M_\star$ from each yield set, assuming a mean Fe yield per SN Ia of $\overline m_{\rm Fe}^{\rm Ia}=0.7$ M$_\odot$ \citep{mazzali_common_2007,howell_effect_2009} \todo{(are there more recent citations?)}. The rate for the {\tt yZ1} yield set is slightly higher than the volumetric rate of $N_{\rm Ia}/M_\star=(1.3\pm0.1)\times10^{-3}\,{\rm M}_\odot^{-1}$ reported by \citet{maoz_star_2017}, but is consistent with their measurement of $N_{\rm Ia}/M_\star=(1.6\pm0.3)\times10^{-3}\,{\rm M}_\odot^{-1}$ for field galaxies. The rate for the {\tt yZ2} yield set is consistent with the measurement of $N_{\rm Ia}/M_\star=(2.2\pm1.0)\times10^{-3}\,{\rm M}_\odot^{-1}$ by \citet{maoz_type-ia_2012}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/mdf_evolution.pdf}
    \caption{Evolution of the MDF across the Galactic disk. Each panel presents the normalized stellar [Fe/H] distribution within a {2 kpc}-wide annulus, color-coded by the age bin. Rows (a) and (b) present the distributions from multi-zone GCE models, while row (c) presents the distributions from APOGEE DR17 with ages derived from [C/N] abundances (see Section \ref{sec:observational-sample}). Distributions are restricted to $0\leq|z|<0.5\,{\rm kpc}$ and boxcar-smoothed with a width of {0.1 dex} in all panels.}
    \label{fig:mdf-evolution}
\end{figure*}

@@

\begin{itemize}
    \item The ``turn-over'' in the evolution of [O/Fe] following the second infall produces a low-$\alpha$ sequence with a fundamentally different abundance structure than observed. A low yield set ($y/Z_\odot=1$) coupled with lower outflows, or pre-enrichment of the infalling gas, can bring the stellar [O/Fe] distributions more in line with the data.
    \item The large quantity of pristine gas accreted in the Solar neighborhood during the second infall phase rapidly dilutes the ISM metallicity by $\sim0.5$ dex. Models with a low yield set ($y/Z_\odot=1$) remain at sub-Solar metallicity until the present day, in stark contrast to the observed local age--metallicity relation. This issue can be mitigated with higher yields and outflows, but not by pre-enriched infall or radial migration.
    \item Our models predict that the MDF evolves to higher metallicity over time throughout the disk. This contrasts with the APOGEE data, which show very little change in the mode over the past $\sim6-8$ Gyr.
\end{itemize}

>>>>

We increased $y_{\rm Fe}^{\rm Ia}$ by a factor of 30\% and 10\% for the {\tt yZ1} and {\tt yZ2} yield sets, respectively, to ensure our models could reach ${\rm [O/Fe]}\approx0.0$ by the end time. The final row of Table \ref{tab:yields} reports the integrated SN Ia rate $N_{\rm Ia}/M_\star$ from each yield set, assuming a mean Fe yield per SN Ia of $\overline m_{\rm Fe}^{\rm Ia}=0.7$ M$_\odot$ \citep{mazzali_common_2007,howell_effect_2009}. The rate for the {\tt yZ1} yield set is slightly higher than the volumetric rate of $N_{\rm Ia}/M_\star=(1.3\pm0.1)\times10^{-3}\,{\rm M}_\odot^{-1}$ reported by \citet{maoz_star_2017}, but is consistent with their measurement of $N_{\rm Ia}/M_\star=(1.6\pm0.3)\times10^{-3}\,{\rm M}_\odot^{-1}$ for field galaxies. The rate for the {\tt yZ2} yield set is consistent with the measurement of $N_{\rm Ia}/M_\star=(2.2\pm1.0)\times10^{-3}\,{\rm M}_\odot^{-1}$ by \citet{maoz_type-ia_2012}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/mdf_evolution.pdf}
    \caption{Evolution of the MDF across the Galactic disk. Each panel presents the normalized stellar [Fe/H] distribution within a {2 kpc}-wide annulus, color-coded by the age bin. Rows (a) and (b) present the distributions from multi-zone GCE models, while row (c) presents the distributions from APOGEE DR17 with ages derived from [C/N] abundances (see Section \ref{sec:observational-sample}). The vertical dotted lines in row (c) mark the mode of the distribution in the $0-2\,{\rm kpc}$ age bin. Distributions are restricted to $0\leq|z|<0.5\,{\rm kpc}$ and boxcar-smoothed with a width of {0.1 dex} in all panels.}
    \label{fig:mdf-evolution}
\end{figure*}

The discrepancies between the predicted and observed abundance evolution discussed in Section \ref{sec:age-abundance} are also seen across the disk. Figure \ref{fig:mdf-evolution} shows the evolution of the MDF with age across five radial bins. We select the $y/Z_\odot=1$ and $y/Z_\odot=2$ models with pre-enriched infall due to the relative success of these models in the previous sections. We use the [C/N]-derived age estimates for the APOGEE sample due to the larger sample size in the most distant regions of the disk; we limit the comparison to ages $\leq10$ Gyr because of large systematic uncertainties for the oldest stars, as discussed in Section \ref{sec:observational-sample}. The two-infall model predicts a very broad MDF for ages $>10$ Gyr because of the rapid enrichment of the ISM during the first infall epoch.

Figure \ref{fig:mdf-evolution} shows that even with pre-enriched infall, both of our models predict that the MDF shifts to the left going from younger to older populations. On the other hand, the data show that the MDF broadens with age, which we attribute to radial migration, but its peak does not shift substantially over the past $\sim6-8$ Gyr. While the mode of the MDF shifts depending on the radial bin, its evolution is remarkably consistent over the entire disk, in line with the equilibrium scenario of \citet{johnson_milky_2024}.

On the other hand, the model predictions for the MDF in the $8-10\,{\rm Gyr}$ age bin are strikingly similar to the APOGEE MDF. The two-infall model predicts a bimodal distribution: the metal-rich component comprises stars formed during the last Gyr of the first infall epoch, and the metal-poor component is stars formed during the first Gyr of the second infall epoch (post-dilution). The data show this same bimodal structure in all but the outermost radial bin, with peaks at ${\rm [Fe/H]}\approx-0.3$ and $+0.3$. While the location of the model peaks depends on $R_{\rm gal}$, in the data they are quite consistent from the innermost to the outermost bin.

@@

\begin{itemize}
    \item The ``turn-over'' in the evolution of [O/Fe] following the second infall produces a low-$\alpha$ sequence with a fundamentally different abundance structure than observed. A low yield set ($y/Z_\odot=1$) coupled with lower outflows, or pre-enrichment of the infalling gas, can bring the stellar [O/Fe] distributions more in line with the data.
    \item The large quantity of pristine gas accreted in the Solar neighborhood during the second infall phase rapidly dilutes the ISM metallicity by $\sim0.5$ dex. Models with a low yield set ($y/Z_\odot=1$) remain at sub-Solar metallicity until the present day, in stark contrast to the observed local age--metallicity relation. This issue can be mitigated with higher yields and outflows, but not by pre-enriched infall or radial migration.
    \item Our models predict that the MDF evolves to higher metallicity over time throughout the disk. This contrasts with the APOGEE data, which show very little change in the mode over the past $\sim6-8$ Gyr.
    \item The equilibrium scenario of chemical evolution, if correct, places stricter limits on the two-infall model than other evolutionary models.
\end{itemize}
